Alex's Notes and Text Fragments for the JSS Manuscript

### 2024-03-15 paper/article.Rmd
trim code to 76 chrs

Turning to prediction tasks with spatiotemporal data, various spatial, temporal, or spatiotemporal partitioning strategies are being used, depending on the specific study objectives.
While the former two ignore the temporal and spatial dimension of the data, respectively, it has also been proposed to leave out random subsets of locations and time points [@meyer2018] or spatiotemporal clusters [@cluto].
Details of these and other implementations are outlined in the respective subsections of Section \ref{sec:implementation}.

@@

autoplot(rsmp_buffer,
  size = 0.8, task = task, fold_id = 1, show_omitted = TRUE
) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(
      size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8
    )
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 1.5)))
```

@@

autoplot(rsmp_disc,
  size = 0.8, task = task, fold_id = 1, show_omitted = TRUE
) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(
      size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8
    )
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 1.5)))
```

@@

autoplot(rsmp_coords,
  size = 0.8, fold_id = 1, task = task
) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(
      size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8
    )
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 1.5)))
```

@@

autoplot(rsmp_tiles,
  size = 0.8, fold_id = 1, task = task
) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(
      size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8
    )
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 1.5)))
```

@@

autoplot(rsmp_custom, size = 0.8, task = task, fold_id = 1) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(
      size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8
    )
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 1.5)))
```

@@

autoplot(rsmp_block_random,
  size = 0.8, fold_id = 1, task = task,
  show_blocks = TRUE, show_labels = TRUE, label_size = 4
) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(
      size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8
    )
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 1.5)))
```

@@

autoplot(rsmp_block_systematic,
  size = 0.8, fold_id = 1, task = task,
  show_blocks = TRUE, show_labels = TRUE, label_size = 4
) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    text = ggplot2::element_text(size = 15),
    plot.title = ggtext::element_textbox(
      size = 12,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8
    )
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 1.5)))
```

@@

```{r cv-eval-2, fig.cap='Cross-Validation at the block level including predefined groups from package \\pkg{mlr3} (method `"cv"`). A factor variable is used to define the grouping. Each class is either assigned to the test or training set.', out.width="60%", fig.pos="ht", echo=FALSE}
autoplot(rsmp_cv_group, size = 0.8, task = task_cv, fold_id = 1) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(
      size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8
    )
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 1.5)))
```

@@

plot_env_single = autoplot(rsmp_env,
  size = 0.3, fold_id = 1, task = task
) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    plot.title = ggtext::element_textbox(
      size = 8,
      r = ggplot2::unit(4, "pt"),
      height = ggplot2::unit(0.30, "inch"), linewidth = 0.5
    ),
    axis.text = ggplot2::element_text(size = 8.5),
    legend.title = ggtext::element_textbox(size = 10),
    legend.text = ggtext::element_textbox(size = 8),
    legend.key.size = ggplot2::unit(0.4, "cm")
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 0.8)))

plot_env_multi = autoplot(rsmp_env_multi,
  size = 0.3, fold_id = 1, task = task
) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    plot.title = ggtext::element_textbox(
      size = 8,
      r = ggplot2::unit(4, "pt"),
      height = ggplot2::unit(0.30, "inch"), linewidth = 0.5
    ),
    axis.text = ggplot2::element_text(size = 8.5),
    legend.title = ggtext::element_textbox(size = 10),
    legend.text = ggtext::element_textbox(size = 8),
    legend.key.size = ggplot2::unit(0.4, "cm")
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 0.8)))

@@

suppressMessages(autoplot(rsmp_knndm,
  size = 0.8, fold_id = 1, task = task, label_size = 4
) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(
      size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8
    )
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 1.5))))
```

@@

```{r vis-spcv-eval, fig.cap="Spatial leave-one-block-out partitioning using coordinate-based clustering to create roughly equally sized polygonal blocks. Due to space limitations only the first two folds of the first repetition are shown.", fig.pos="ht", out.width="100%", echo=FALSE}
autoplot(rsmp_sp, task, fold_id = 1:2, size = 0.3) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) *
  ggplot2::theme(
    plot.title = ggtext::element_textbox(
      size = 8,
      r = ggplot2::unit(4, "pt"),
      height = ggplot2::unit(0.30, "inch"), linewidth = 0.5
    ),
    axis.text = ggplot2::element_text(size = 8.5),
    legend.title = ggtext::element_textbox(size = 10),
    legend.text = ggtext::element_textbox(size = 8),
    legend.key.size = ggplot2::unit(0.4, "cm")
  ) *
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 0.8)))
```

@@

```{r vis-nspcv-eval, fig.cap="Random (non-spatial) four-fold CV partitioning. Only the first two folds of the first repetition are shown.", out.width="100%", fig.pos="ht", echo=FALSE}
autoplot(rsmp_nsp, task, fold_id = 1:2, size = 0.3) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) *
  ggplot2::theme(
    plot.title = ggtext::element_textbox(
      size = 8,
      r = ggplot2::unit(4, "pt"),
      height = ggplot2::unit(0.30, "inch"), linewidth = 0.5
    ),
    axis.text = ggplot2::element_text(size = 8.5),
    legend.title = ggtext::element_textbox(size = 10),
    legend.text = ggtext::element_textbox(size = 8),
    legend.key.size = ggplot2::unit(0.4, "cm")
  ) *
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 0.8)))
```

>>>>

Turning to prediction tasks with spatiotemporal data, various spatial, temporal, or spatiotemporal partitioning strategies are being used, depending on the specific study objectives.
While the former two ignore the temporal and spatial dimension of the data, respectively, it has also been proposed to leave out random subsets of locations and time points [@meyer2018] or spatiotemporal clusters [@cluto].
Details of these and other implementations are outlined in Section \ref{sec:implementation}.

@@

autoplot(rsmp_buffer,
  size = 0.8, task = task, fold_id = 1, show_omitted = TRUE
) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(
      size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8
    )
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(
    override.aes = list(size = 1.5)))
```

@@

autoplot(rsmp_disc,
  size = 0.8, task = task, fold_id = 1, show_omitted = TRUE
) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(
      size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8
    )
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(
    override.aes = list(size = 1.5)))
```

@@

autoplot(rsmp_coords,
  size = 0.8, fold_id = 1, task = task
) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(
      size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8
    )
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(
    override.aes = list(size = 1.5)))
```

@@

autoplot(rsmp_tiles,
  size = 0.8, fold_id = 1, task = task
) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(
      size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8
    )
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(
    override.aes = list(size = 1.5)))
```

@@

autoplot(rsmp_custom, size = 0.8, task = task, fold_id = 1) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(
      size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8
    )
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(
    override.aes = list(size = 1.5)))
```

@@

autoplot(rsmp_block_random,
  size = 0.8, fold_id = 1, task = task,
  show_blocks = TRUE, show_labels = TRUE, label_size = 4
) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(
      size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8
    )
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(
    override.aes = list(size = 1.5)))
```

@@

autoplot(rsmp_block_systematic,
  size = 0.8, fold_id = 1, task = task,
  show_blocks = TRUE, show_labels = TRUE, label_size = 4
) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    text = ggplot2::element_text(size = 15),
    plot.title = ggtext::element_textbox(
      size = 12,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8
    )
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(
    override.aes = list(size = 1.5)))
```

@@

```{r cv-eval-2, fig.cap='Cross-Validation at the block level including predefined groups from package \\pkg{mlr3} (method `"cv"`). A factor variable is used to define the grouping. Each class is either assigned to the test or training set.', out.width="60%", fig.pos="ht", echo=FALSE}
autoplot(rsmp_cv_group, size = 0.8, task = task_cv, fold_id = 1) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(
      size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8
    )
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(
    override.aes = list(size = 1.5)))
```

@@

plot_env_single = autoplot(rsmp_env,
  size = 0.3, fold_id = 1, task = task
) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    plot.title = ggtext::element_textbox(
      size = 8,
      r = ggplot2::unit(4, "pt"),
      height = ggplot2::unit(0.30, "inch"), linewidth = 0.5
    ),
    axis.text = ggplot2::element_text(size = 8.5),
    legend.title = ggtext::element_textbox(size = 10),
    legend.text = ggtext::element_textbox(size = 8),
    legend.key.size = ggplot2::unit(0.4, "cm")
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(
    override.aes = list(size = 0.8)))

plot_env_multi = autoplot(rsmp_env_multi,
  size = 0.3, fold_id = 1, task = task
) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    plot.title = ggtext::element_textbox(
      size = 8,
      r = ggplot2::unit(4, "pt"),
      height = ggplot2::unit(0.30, "inch"), linewidth = 0.5
    ),
    axis.text = ggplot2::element_text(size = 8.5),
    legend.title = ggtext::element_textbox(size = 10),
    legend.text = ggtext::element_textbox(size = 8),
    legend.key.size = ggplot2::unit(0.4, "cm")
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(
    override.aes = list(size = 0.8)))

@@

suppressMessages(autoplot(rsmp_knndm,
  size = 0.8, fold_id = 1, task = task, label_size = 4
) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(
      size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8
    )
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(
    override.aes = list(size = 1.5))))
```

@@

```{r vis-spcv-eval, fig.cap="Spatial leave-one-block-out partitioning using coordinate-based clustering to create roughly equally sized polygonal blocks. Due to space limitations only the first two folds of the first repetition are shown.", fig.pos="ht", out.width="100%", echo=FALSE}
autoplot(rsmp_sp, task, fold_id = 1:2, size = 0.3) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) *
  ggplot2::theme(
    plot.title = ggtext::element_textbox(
      size = 8,
      r = ggplot2::unit(4, "pt"),
      height = ggplot2::unit(0.30, "inch"), linewidth = 0.5
    ),
    axis.text = ggplot2::element_text(size = 8.5),
    legend.title = ggtext::element_textbox(size = 10),
    legend.text = ggtext::element_textbox(size = 8),
    legend.key.size = ggplot2::unit(0.4, "cm")
  ) *
  ggplot2::guides(colour = ggplot2::guide_legend(
    override.aes = list(size = 0.8)))
```

@@

```{r vis-nspcv-eval, fig.cap="Random (non-spatial) four-fold CV partitioning. Only the first two folds of the first repetition are shown.", out.width="100%", fig.pos="ht", echo=FALSE}
autoplot(rsmp_nsp, task, fold_id = 1:2, size = 0.3) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) *
  ggplot2::theme(
    plot.title = ggtext::element_textbox(
      size = 8,
      r = ggplot2::unit(4, "pt"),
      height = ggplot2::unit(0.30, "inch"), linewidth = 0.5
    ),
    axis.text = ggplot2::element_text(size = 8.5),
    legend.title = ggtext::element_textbox(size = 10),
    legend.text = ggtext::element_textbox(size = 8),
    legend.key.size = ggplot2::unit(0.4, "cm")
  ) *
  ggplot2::guides(colour = ggplot2::guide_legend(
    override.aes = list(size = 0.8)))
```
